---
execute:
  cache: false
---

# Regions and their Countries {#sec-02-countries-in-regions}

```{r}
#| label: setup
#| results: hold
#| include: false

base::source(file = paste0(here::here(), "/R/helper.R"))

ggplot2::theme_set(ggplot2::theme_bw())
```

## Objectives {.unnumbered}

::::: my-objectives
::: my-objectives-header
Country and Regions: Classifications for countries
:::

::: my-objectives-container
1.  Understanding the country classification used by my `r glossary("WHR")` data sources (@sec-02-whr-classification).
2.  Inspecting different approaches to classify countries by
    international organizations:
    - World Bank and 
    - United Nations Statistics Division
3.  Modifying resp. adding country classifications to the WHR data where necessary
    so that it conforms to the internationally
    recognized and approved systems.

I aim to compare different aspects of 
countries. I want, for instance, to know how well Austria is doing
compared to other European countries, the other member states of the
European Union, or other OECD countries. It is, therefore, vital to have
a consistent categorization system with different grouping schemes.
:::
:::::

## Country groupings in WHR {#sec-02-whr-classification}

Some of the WHR years (WHR reports 2013, 2015, 2020, and 2021) have a regional
grouping incorporated. But this grouping is not included in the (new) dataset 2024 from the WHR report 2025.

### Classification of WHR 2020

I will display the regional grouping of the WHR data with the example from the year 2020 (WHR report 2021).

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-02-whr-classification-2020}
: WHR data 2020 classification
::::::
:::
::::{.my-r-code-container}
::: {#lst-whr-classification-2020}
```{r}
#| label: whr-classification-2020


(
  df_dt_whr_2020 <-  base::readRDS("data/whr/raw/whr_raw_2021.rds") |> 
      dplyr::select(`Country name`, `Regional indicator`) |> 
      dplyr::nest_by(`Regional indicator`) |> 
      dplyr::mutate(data = as.vector(data)) |>
      dplyr::mutate(data = stringr::str_c(data, collapse = "; ")) |>
      dplyr::mutate(data = paste(data, ";")) |> 
      dplyr::mutate(N = lengths(gregexpr(";", data))) |> 
      dplyr::rename(Country = data) |> 
      DT::datatable(class = 'cell-border compact stripe', 
                options = list(
                  pageLength = 25,
                  lengthMenu = c(5, 10, 15, 20, 25, 50)
                  )
            )
)
```
Result of the WHR classification system for WHR data 2020 used in the WHR report 2021.
:::




------------------------------------------------------------------------

There are `r base::length(df_dt_whr_2020$x$data$N)` different regional indicators. The datasets for 2013, 2015, 2020 and 2021 use all the same
classification scheme with **`r base::sum(df_dt_whr_2020$x$data$N)` countries in
`r base::length(df_dt_whr_2020$x$data$N)` regions**.

`r sum(df_dt_whr_2020$x$data$N)` are by far not all countries of the world. Their complete number is about 195 with some insecurities about Holy See (Vatican), the State of Palestine, Taiwan and Kosovo. (Compare: [How Many Countries Are There In The World?](https://www.worldatlas.com/geography/how-many-countries-are-there-in-the-world.html)) The reason for this lower number is simple: For only those `r sum(df_dt_whr_2020$x$data$N)` countries are subjective well-being data in the study year 2020 available.

::::
:::::

### `class_scheme()` function

As I am going to list several classification variants it pays the effort to develop a function for the repetitive task.

::::::: my-r-code
:::: my-r-code-header
::: {#cnj-show-class-scheme}
: Function `class_scheme()` for showing classification schemes
:::
::::

:::: my-r-code-container
::: {#lst-show-class-scheme}
```{r}
#| label: function-class-scheme
#| code-fold: show

class_scheme <- function(df, sel1, sel2) {
    ## df = dataframe to show
    ## sel1 = name of the first column (country names) to select
    ## sel2 = name of the column with the regional indicator
  df |> 
        dplyr::select(!!sel1, !!sel2) |> 
        dplyr::nest_by(!!sel2) |> 
        dplyr::mutate(data = as.vector(data)) |>
        dplyr::mutate(data = stringr::str_c(data, collapse = "; ")) |>
        dplyr::mutate(data = paste(data, ";")) |> 
        dplyr::mutate(N = lengths(gregexpr(";", data))) |> 
        dplyr::rename(Country = data) |> 
        dplyr::arrange(!!sel2) |> 
        DT::datatable(class = 'cell-border compact stripe', 
            options = list(
              pageLength = 25,
              lengthMenu = c(5, 10, 15, 20, 25, 50)
              )
        )
}



```

Function `class_scheme()` for showing results of a classification system
:::

Here I am using complex code lines. Using {**dplyr**} programming code in functions needs some special consideration. I have learned the details from "Bang Bang – How to
program with dplyr" [@berroth-2019].
::::
:::::::

### WHR 2020 with `class_scheme()` function

As the `class_scheme()` function is now in place, I can display with this function the different grouping
schemes. At first I will try it out with the WHR data from the 2021 report:

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-02-whr-grouping-schema}
: Classification of the WHR data
::::::
:::
::::{.my-r-code-container}
```{r}
#| label: whr-grouping-schema

df_whr <-  base::readRDS(
    paste0(here::here(), "/data/whr/raw/whr_raw_2021.rds"))
(
    whr_class <- class_scheme(
            df = df_whr,
            sel1 = rlang::quo(`Country name`),
            sel2 = rlang::quo(`Regional indicator`)
            )
)
```

::::
:::::


It worked! I got the same result as in @lst-whr-classification-2020.


## Official classifications

There are already different classification systems in place:
International organizations (e.g., [World
Bank](https://datahelpdesk.worldbank.org/knowledgebase/articles/906519),
[United Nations](https://unstats.un.org/unsd/methodology/m49/)) have
developed them with several grouping variants.

I will look into these two official classifications schemes of World Bank and United Nations and apply the following procedure:

:::::{.my-procedure}
:::{.my-procedure-header}
:::::: {#prp-country-class}
: Understand structure and content of the official classifications schemata
::::::
:::
::::{.my-procedure-container}
1. Create a directory for storing the different country classification files (see @sec-02-create-data-dirs).
2. Download classification files and store them for faster access as R objects with `r glossary("rds")` format  (see @sec-02-wb-download and @sec-02-unsd-download).
3. Inspect the data classification files of World Bank (@sec-02-inspect-wb) and of the United Nations (@sec-02-inspect-unsd) in detail.

::::
:::::

### Create data directories {#sec-02-create-data-dirs}

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-create-class-dirs}
: Create folders for country classification files
::::::
:::
::::{.my-r-code-container}



```{r}
#| label: create-class-dirs
#| code-fold: show

my_create_folder(base::paste0(here::here(), "/data/"))
my_create_folder(base::paste0(here::here(), "/data/country-class"))
my_create_folder(base::paste0(here::here(), "/data/country-class/wb"))
my_create_folder(base::paste0(here::here(), "/data/country-class/unsd"))
my_create_folder(base::paste0(here::here(), "/data/country-class/wb/excel"))
my_create_folder(base::paste0(here::here(), "/data/country-class/wb/csv"))
my_create_folder(base::paste0(here::here(), "/data/country-class/wb/rds"))
my_create_folder(base::paste0(here::here(), "/data/country-class/unsd/excel"))
my_create_folder(base::paste0(here::here(), "/data/country-class/unsd/csv"))
my_create_folder(base::paste0(here::here(), "/data/country-class/unsd/rds"))
```

<center>(*For this R code chunk is no output available*)</center>


::::
:::::

### Download classification files {#sec-02-download-class-files}

#### World Bank {#sec-02-wb-download}

The World Bank Classification can be downloaded from [How does the World Bank classify countries?](https://datahelpdesk.worldbank.org/knowledgebase/articles/378834-how-does-the-world-bank-classify-countries). Near the bottom of the page you can see the line "Download an Excel file of historical classifications by income.", providing a link with the word "Download". The downloaded file `CLASS.xlsx` does *not* contain a historical classification by income but the general classification system of the last available year (2023).  

Yes, there is another Excel file `OGHIST.xslx` with the historical cutoffs
for incomes and lending categories, dating from 1987 to 2023. But the download link for this file is located at another web page: [World Bank Country and Lending Groups](https://datahelpdesk.worldbank.org/knowledgebase/articles/906519). On this page you will also find the updates for the cutoffs for countries `r glossary("GNIx", "GNI")` income per capita which is important for the lending eligibility of countries. [World Bank country classifications by income level for
2024-2025](https://blogs.worldbank.org/en/opendata/world-bank-country-classifications-by-income-level-for-2024-2025) has the current updated values and changes over the last year.

The file `CLASS.xlsx` I am interested here consists of three sheets. 

  - 1. "List of Economies"
  - 2. "compositions" and
  - 3. "Notes"

I will download the original Excel file with all it sheets and save programmatically
  
  - the Excel file with all its sheet
  - CSV snapshots of all sheets (file extension = `.csv`) and
  - R objects of all sheets (file extension = `.rds`)
  
The CSV snapshots support reproducibility because it stores the proprietary Excel file in in a tool-agnostic, future-proof format. I am using code inspired by the vignette/article [readxl Workflows](https://readxl.tidyverse.org/articles/readxl-workflows.html) 


:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-02-wb-class}
: Download the World Bank CLASS Excel file
::::::
:::
::::{.my-r-code-container}

<center>**Run this code chunk manually if the file still needs to be downloaded.**</center>


```{r}
#| label: wb-class
#| code-fold: show
#| eval: false

url_excel = "https://datacatalogfiles.worldbank.org/ddh-published/0037712/DR0090755/CLASS.xlsx"
path_wb_excel <- base::paste0(here::here(), 
            "/data/country-class/wb/excel/wb-class.xlsx")
path_wb_csv <- base::paste0(here::here(), 
            "/data/country-class/wb/csv/")
path_wb_rds <- base::paste0(here::here(), 
            "/data/country-class/wb/rds/")

## download wb-class file ##############
downloader::download(
    url = url_excel,
    destfile = path_wb_excel
)

## from readxl workflow article ##############
## includes also my_excel_as_csv_and_rds() function 
path_wb_excel |> 
  readxl::excel_sheets()  |> 
  rlang::set_names()  |>  
  purrr::map(my_excel_as_csv_and_rds, 
             path_excel = path_wb_excel, 
             path_csv = path_wb_csv,
             path_rds = path_wb_rds
             ) 
```

::::
:::::

#### UNSD-M49 {#sec-02-unsd-download}

Another more detailed classification system expressively developed for
statistical purposes is developed by the United Nations Statistics
Division `r glossary("UNSD")` using the `r glossary("M49")` methodology.

The result is called [Standard country or area codes for statistical use
(M49)](https://unstats.un.org/unsd/methodology/m49/) and can be
downloaded manually in different languages and formats (Copy into the
clipboard, Excel or `r glossary("CSV")` from the [Overview
page](https://unstats.un.org/unsd/methodology/m49/overview/). On the page "Overview" is
no URL for an R script available, because triggering one of the buttons
copies or downloads the data with the help of Javascript. So I had to download the file manually or to find another location where I could download it programmatically. 

I found with the `r glossary("OMNIKA")` DataStore [United Nations M49 Region Codes](https://omnika.org/datastore/datasets/un-m49-region-codes) an [external source for the UNSD-M49 country classification](https://github.com/omnika-datastore/unsd-m49-standard-area-codes). For security reason I checked the two files with `base::all.equal()` to determine if those two files are identical. Yes, they are!

The UNSD M40 standard area codes are stored as Excel and CSV files. I download for reproducibility reason the CSV file.

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-02-UNSD-M49}
: Download the UNSD-M49 CSV file and create an R object (".rds")
::::::
:::
::::{.my-r-code-container}

<center>**Run this code chunk manually if the file still needs to be downloaded.**</center>

```{r}
#| label: unsd-class
#| code-fold: show
#| eval: false

## download unsd-m49 file ############
url_unsd_csv <- "https://github.com/omnika-datastore/unsd-m49-standard-area-codes/raw/refs/heads/main/2022-09-24__CSV_UNSD_M49.csv"
path_unsd_csv <- base::paste0(here::here(), 
     "/data/country-class/unsd/csv/2022-09-24__CSV_UNSD_M49.csv")

downloader::download(
    url = url_unsd_csv,
    destfile = path_unsd_csv
)


## create R object ###############
unsd_class <- 
  readr::read_delim(
    file = path_unsd_csv,
    delim = ";"
  )


## save as .rds file ################
my_save_data_file(
  "country-class/unsd/rds", 
  unsd_class, 
  "unsd_class.rds")



```

<center>(*For this R code chunk is no output available*)</center>

::::
:::::





### Inspect classification files {#sec-02-inspect-class-files}

To get an detailed understanding of the data structures I will provide the following outputs:

1. A summary statistics with `skimr::skim()` followed by inspection of the first data with `dplyr::glimpse()`.
2. Several detailed outputs of the classifications categories (regions) and their elements (countries) in different code chunks (tabs).

#### World Bank {#sec-02-inspect-wb}

::: {.my-code-collection}
:::: {.my-code-collection-header}
::::: {.my-code-collection-icon}
:::::
:::::: {#exm-02-inspect-wb-class-files}
: Inspect the structure of the World Bank classification
::::::

::::

::::{.my-code-collection-container}

::: {.panel-tabset}

###### WB `economies`

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-02-inspect-wb-sheet1}
: Inspect sheet `List of Economies` of the World Bank classification file
::::::
:::
::::{.my-r-code-container}
```{r}
#| label: inspect-wb-sheet1
#| results: hold

wb_class_economies <- base::readRDS(
  "data/country-class/wb/rds/wb-class-List of economies.rds")
glue::glue("******************* Using skimr::skim() ***************************")
skimr::skim(wb_class_economies)
glue::glue("")
glue::glue("****************** Using dplyr::glimpse() *************************")
dplyr::glimpse(wb_class_economies)

```


::::
:::::


###### WB `compositions`

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-02-inspect-wb-sheet2}
: Inspect sheet `compositions` of the World Bank classification file
::::::
:::
::::{.my-r-code-container}
```{r}
#| label: inspect-wb-sheet2
#| results: hold

wb_class_compositions <- base::readRDS(
  "data/country-class/wb/rds/wb-class-compositions.rds")
glue::glue("******************* Using skimr::skim() ***************************")
skimr::skim(wb_class_compositions)
glue::glue("")
glue::glue("****************** Using dplyr::glimpse() *************************")
dplyr::glimpse(wb_class_compositions)


```



::::
:::::

###### WB standard

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-02-wb-class-standard}
: Pre-defined standard categorization
::::::
:::
::::{.my-r-code-container}
```{r}
#| label: wb-class-standard


df_wb_standard <-  base::readRDS(
  "data/country-class/wb/rds/wb-class-List of economies.rds") |> 
    dplyr::slice(1:218)



(
    wb_class_standard <- class_scheme(
            df = df_wb_standard,
            sel1 = rlang::quo(`Economy`),
            sel2 = rlang::quo(`Region`)
            )
)
```

***

`Region` is a coarse classification scheme with only **`r length(wb_class_standard$x$data$N)` regions formed by `r sum(wb_class_standard$x$data$N)` countries**.

::::
:::::

###### WB All

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-02-wb-class-all}
: All provided groups, regional, economical and political
::::::
:::
::::{.my-r-code-container}
```{r}
#| label: wb-class-all

df_wb_all <-  base::readRDS(
  "data/country-class/wb/rds/wb-class-compositions.rds")


(
    wb_class_all <- class_scheme(
            df = df_wb_all,
            sel1 = rlang::quo(`WB_Country_Name`),
            sel2 = rlang::quo(`WB_Group_Name`)
            )
)
```

***

`WB_Group_Name` in the "compositions" file contains all available groups. They are not restricted to regional groups because they are formed by economical and political criteria as well. There is no 1:1 match, because almost all countries belong to two or more groups. There are **`r length(wb_class_all$x$data$N)` groups with a total of `r sum(wb_class_all$x$data$N)` elements**.

::::
:::::

###### Region1

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-02-wb-class-regional}
: Groups formed by regional criteria (without the redundant `World` region)
::::::
:::
::::{.my-r-code-container}
```{r}
#| label: wb-class-regional

str_reg <- c("AFE", "AFW", "ARB", "CSS", "CEB",
             "EAS", "ECS", "LCN", "MEA", "NAC",
             "OSS", "PSS", "SST", "SAS", "SSF")

df_wb_reg <-  base::readRDS(
  "data/country-class/wb/rds/wb-class-compositions.rds") |> 
  dplyr::filter(WB_Group_Code %in% str_reg)

(
    wb_class_reg <- class_scheme(
            df = df_wb_reg,
            sel1 = rlang::quo(`WB_Country_Name`),
            sel2 = rlang::quo(`WB_Group_Name`)
            )
)
  
```

***

Browsing through the `composition` data I have defined 15 `WB_GROUP_CODE`s as regional codes. These regional classification criteria results per definition to **`r length(wb_class_reg$x$data$N)` regions containing `r sum(wb_class_reg$x$data$N)` countries**.

::::
:::::

###### Region2

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-02-wb-class-regional}
: Groups formed by regional criteria (without the redundant `World` region)
::::::
:::
::::{.my-r-code-container}
```{r}
#| label: wb-class-regional2

str_reg2 <- c("AFE", "AFW", "ARB", "CEB",
             "EAS", "ECS", "LCN", "MEA", "NAC",
             "SAS", "SSF")

df_wb_reg2 <-  base::readRDS(
  "data/country-class/wb/rds/wb-class-compositions.rds") |> 
  dplyr::filter(WB_Group_Code %in% str_reg2)

(
    wb_class_reg2 <- class_scheme(
            df = df_wb_reg2,
            sel1 = rlang::quo(`WB_Country_Name`),
            sel2 = rlang::quo(`WB_Group_Name`)
            )
)
  
```

***

Browsing through the `composition` data I have declassified all small states for an alternative regional group. These regional classification criteria results to **`r length(wb_class_reg2$x$data$N)` regions containing `r sum(wb_class_reg2$x$data$N)` countries**.

::::
:::::



:::

::::
:::::

##### Description of the World Bank tabs

1. **WB economies** displays the "List of Economies" and has five columns:
    -   `Economy` with the country names (2-219) and regional names
        (221-268)
    -   `Code` with the ISO alpha3 codes for countries (2-219) and for
        the regional names (221-268)
    -   `Region` with seven different regional names:
        -   East Asia and Pacific,
        -   Europe and Central Asia,
        -   Latin America & the Caribbean,
        -   Middle East and North Africa,
        -   North America
        -   South Asia and
        -   Sub-Saharan Africa
    -   `Income group` with four groups: Low income, Lower middle
        income, Higher middle income, and High income.
    -   `Lending category` with three groups: `IBRD`, `Blend`, and
        `IDA`.
2.  **WB compositions** has four columns: `WB_Group_Code`, `WB_Group_Name`,
    `WB_Country_Code`, `WB_Country_Name`. The 2084 rows are combinations
    of the regional and income group with their ISO alpha 3 codes and
    country names.
3. **WB Standard** shows the World Bank seven standard regional groups with their countries. The 218 countries  involved in the taxonomy of the World Bank consists of all member countries of the World Bank (189) and other economies with populations of more than 30,000 (29).
4. **WB All** includes the seven regions from the "WB Standard" tab but much more. But it is important to note that there is no alternative *regional* structure that comprises systematically all countries of the world --- the overall category "World" obviously excluded. 
    - Five of the seven regional groups of "WB Standard" are also clustered without high income countries. 
    - There are six other regional subcategories: "Arab World", "Caribbean small states", "Central Europe and Baltics", "Other small states", "Pacific island small states", "Small states". 
    - Additionally there are some political groups like European Union, OECD and 
    - several economical classification like "Euro area", 
    - different combinations of the four income groups and 
    different combinations of the three lending statuses.
5. **Region 1** includes `WB_Group_Code` from the `composition` data and results into 15 regions containing 379 countries.
6. **Region 2** includes again the `WB_Group_Code` from the `composition` data but has excluded all small states for an alternative regional group. These regional classification criteria results to 11 regions containing 299 countries.
    
**More details**

The cut off limits for the income groups are: 

- low income, $1,145 or less; 
- lower middle income, $1,146  to $4,515; 
- upper middle income, $4,516 to $14,005; and 
- high income, more than $14,005. 

The effective operational cutoff for `r glossary("IDAx", "IDA")` eligibility is $1,335 or less. The three lending categories and their relation to each other are:

> `r glossary("IDAx", "IDA")` countries are those that lack the financial ability to borrow from `r glossary("IBRD")`. IDA credits are deeply concessional—interest-free loans and grants for programs aimed at boosting economic growth and improving living conditions. IBRD loans are non-concessional. `Blend` countries are eligible for IDA credits because of their low per capita incomes but are also eligible for IBRD because they are financially creditworthy.

Three additional remark relating to the `Notes` sheet: 

1. In the `Notes` I found the sentence: "Geographic classifications in this table cover all income levels." But there is a difference of one missing data value more in the `Income group` column compared with the `Region` column (50:49). The reason is that `Venezuela RB` is lacking an income group because it has been temporarily unclassified since July 2021 pending release of revised national accounts statistics. Venezuela, RB was classified as an upper-middle income country until FY21, has been unclassified since then due to the unavailability of data. But it is now again classified as `Upper middle income` (See the World Bank [page about Venezuela, RB](https://archive.doingbusiness.org/en/data/exploreeconomies/venezuela)).

2. The term country, used interchangeably with economy, does not imply political independence but refers to any territory for which authorities report separate social or economic statistics. 

3. What follows is a quote about some details of the income classifications for the 2023 file:

> Set on 1 July 2022 remain in effect until 1 July 2023. Venezuela has been temporarily unclassified since July 2021 pending release of revised national accounts statistics. Argentina, which was temporarily unclassified in July 2016 pending release of revised national accounts statistics, was classified as upper middle income for FY17 as of 29 September 2016 based on alternative conversion factors. Also effective 29 September 2016, Syrian Arab Republic is reclassified from IBRD lending category to IDA-only. On 29 March 2017, new country codes were introduced to align World Bank 3-letter codes with ISO 3-letter codes: Andorra (AND), Dem. Rep. Congo (COD), Isle of Man (IMN), Kosovo (XKX), Romania (ROU), Timor-Leste (TLS), and West Bank and Gaza (PSE). It is to be noted that Venezuela, RB classified as an upper-middle income country until FY21, has been unclassified since then due to the unavailability of data.

##### Summary {#sec-02-wb-summary}

The only missing data in the columns `Economy` and `Code` corresponds to the empty line #220 that separates the country codes from the regional codes. The missing data in the other columns stem from the different structure of the second part (starting with row #221) of the data, which consists only of the two columns 'Economy' and 'Code'. 

Essentially this means that we have in the `wb-class.xlsx` file two different data sets: One for economies and the other one to explicate regional, economical and political grouping codes. In the Excel sheet `compositions` you will find an extended list of all available group names and their three letter codes combined with the country names and their three letter codes. These group names comprise different kinds of regional groups but also names and codes for different combination of country incomes and lending categories.

All these groups may be of interests for analysis of different trends. But the regional (sub)groups of the `compositions` sheet do not add up to the complete number of countries (218). This is in contrast to the different regional groups of the WID database because all their regional groups (region1 = 5, region2 = 18, region4 = 10, and region5 = 8 groups) includes all countries (in this case: 216).

The World Bank file `wb-class.xlsx` classifies all World Bank member countries (189), and all other economies with populations of more than 30,000 (29) in a coarse grid of only seven regions. For operational and analytical purposes, these economies are divided among income groups according to their [gross national income (GNI) per capita](https://datahelpdesk.worldbank.org/knowledgebase/articles/378831-why-use-gni-per-capita-to-classify-economies-into) in 2023, calculated using the [World Bank Atlas method](https://datahelpdesk.worldbank.org/knowledgebase/articles/378832-what-is-the-world-bank-atlas-method). 


#### United Nations {#sec-02-inspect-unsd}



::: {.my-code-collection}
:::: {.my-code-collection-header}
::::: {.my-code-collection-icon}
:::::
:::::: {#exm-02-inspect-unsd}
: Inspect UNSD-M49 geoscheme classification
::::::

::::

::::{.my-code-collection-container}

::: {.panel-tabset}

###### raw

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-02-inspect-unsd-m49}
: Inspect UNSD M49 geoscheme classification
::::::
:::
::::{.my-r-code-container}
```{r}
#| label: inspect-unsd-m49
#| results: hold

unsd_class <- base::readRDS(
  "data/country-class/unsd/rds/unsd_class.rds")
glue::glue("******************* Using skimr::skim() ***************************")
skimr::skim(unsd_class)
glue::glue("")
glue::glue("****************** Using dplyr::glimpse() *************************")
dplyr::glimpse(unsd_class)

```

::::
:::::


###### clean 

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-02-clean-unsd-m49}
: Clean UNSD M49 geoscheme classification
::::::
:::
::::{.my-r-code-container}
:::::{#lst-02-clean-unsd-m49}
```{r}
#| label: clean-unsd-49
#| results: hold

## column renaming vector ########
m49_cols = c(
  region_c = "Region Code", region_n = "Region Name",
  subr_c = "Sub-region Code", subr_n = "Sub-region Name", 
  midr_c = "Intermediate Region Code", midr_n = "Intermediate Region Name",
  country = "Country or Area", m49 = "M49 Code", 
  iso2 = "ISO-alpha2 Code", iso3 = "ISO-alpha3 Code",
  ldc = "Least Developed Countries (LDC)", 
  lldc = "Land Locked Developing Countries (LLDC)", 
  sids = "Small Island Developing States (SIDS)"
  )
  
## clean data ###############################
unsd_class <- base::readRDS(
  "data/country-class/unsd/rds/unsd_class.rds")
unsd_class_clean <- unsd_class |> 
  dplyr::select(-(1:2)) |> 
  dplyr::rename(tidyselect::all_of(m49_cols)) |> 
  dplyr::filter(country != "Antarctica") |> 
  dplyr::mutate(iso2 = base::ifelse(country == "Namibia", "NA", iso2)) |> 
  dplyr::relocate(country, .before = region_c) |> 
  # .x = anonymous function; "x" = value in cols of unsd_class
  dplyr::mutate(dplyr::across(
    ldc:sids, ~ dplyr::if_else(.x == "x", "1", "999", "0") 
    )) |> 
  dplyr::arrange(country)

## save new tibble ##########
my_save_data_file(
  "country-class/unsd/rds",
  unsd_class_clean,
  "unsd_class_clean.rds"
)


## prepare skimmers ##########
my_skim <- skimr::skim_with(
  character = skimr::sfl(
    whitespace = NULL,
    min = NULL,
    max = NULL,
    empty = NULL
    )
)


## display results ##########
unsd_class <- base::readRDS(
  "data/country-class/unsd/rds/unsd_class.rds")
glue::glue("******************* Using skimr::skim() ***************************")
my_skim(unsd_class_clean) |> dplyr::select(-complete_rate)
glue::glue("")
glue::glue("****************** Using dplyr::glimpse() *************************")
dplyr::glimpse(unsd_class_clean)

```

Script for data cleaning of the `unsd_class.rds` file as explained in @prp-02-clean-unsd-m49-data
:::::

::::
:::::

###### Region

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-02-ID-text}
: Display regions of UNSD class scheme
::::::
:::
::::{.my-r-code-container}
```{r}
#| label: unsd-class1

df_unsd <-  base::readRDS(
  "data/country-class/unsd/rds/unsd_class_clean.rds")

(
    unsd_class1 <- class_scheme(
            df = df_unsd,
            sel1 = rlang::quo(`country`),
            sel2 = rlang::quo(`region_n`)
            )
)
```



::::
:::::

`region_n` is a classification scheme for continents with **`r sum(unsd_class1$x$data$N)`
countries in `r length(unsd_class1$x$data$N)` regions**.

###### Sub-region

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-02-sub-region}
: Display sub-regions of UNSD class scheme
::::::
:::
::::{.my-r-code-container}
```{r}
#| label: unsd-class2

df_unsd <-  base::readRDS(
  "data/country-class/unsd/rds/unsd_class_clean.rds")

(
    unsd_class2 <- class_scheme(
            df = df_unsd,
            sel1 = rlang::quo(`country`),
            sel2 = rlang::quo(`subr_n`)
            )
)
```

::::
:::::

`subr_n` is a classification scheme with **`r sum(unsd_class2$x$data$N)`
countries in `r length(unsd_class2$x$data$N)` regions**.

###### Intermediate

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-ID-text}
: Display intermediate regions of UNSD class scheme
::::::
:::
::::{.my-r-code-container}
```{r}
#| label: unsd-class3

df_unsd <-  base::readRDS(
  "data/country-class/unsd/rds/unsd_class_clean.rds")

(
    dt_unsd_class3 <- class_scheme(
            df = df_unsd,
            sel1 = rlang::quo(`country`),
            sel2 = rlang::quo(`midr_n`)
            )
)
```


::::
:::::

`midr_n` is an inconsistent classification scheme with **`r sum(dt_unsd_class3$x$data$N)`
countries in `r length(dt_unsd_class3$x$data$N)` regions**. It has a focus on South- / Central America and Africa. All the other countries are together in a big mixed groups with 140 countries.

###### Intermediate2

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-02-ID-text}
: Display alternative intermediate regions of UNSD class scheme
::::::
:::
::::{.my-r-code-container}
```{r}
#| label: unsd-class4
#| results: hold

unsd_class_clean <-  base::readRDS(
  "data/country-class/unsd/rds/unsd_class_clean.rds")

unsd_class4 <- unsd_class_clean |> 
  dplyr::mutate(midr_n2 = 
         base::ifelse(is.na(midr_n), subr_n, midr_n)
         ) |> 
  dplyr::mutate(midr_c2 = 
         base::ifelse(is.na(midr_c), subr_c, midr_c)
         )


## save intermediate2 data as .rds file ################
my_save_data_file(
  "country-class/unsd/rds", 
  unsd_class4, 
  "unsd_class_finale.rds")


(
    dt_unsd_class4 <- class_scheme(
            df = unsd_class4,
            sel1 = rlang::quo(`country`),
            sel2 = rlang::quo(`midr_n2`)
            )
) 
  
```


::::
:::::

`midr_n2` is a classification scheme with **`r sum(dt_unsd_class4$x$data$N)`
countries in `r length(dt_unsd_class4$x$data$N)` regions**. This grouping divides each continent into different regions and has more plausibility than the `midr_n` classification in tab "Intermediate".



:::

::::
:::::


##### Descriptions of the UNSD-M49 geoscheme classification

What follows is a description if the tabs in @exm-02-inspect-unsd.

**Tab "raw"**: The raw file `unsd_class` has 15 columns as you can also see online from the [Overview
page](https://unstats.un.org/unsd/methodology/m49/overview/). The many missing values (`NAs`) for the categories `r glossary("LDCx", "LDC")`, `r glossary("LLDC")` and `r glossary("SIDS")` are easy explained: These three columns are coded with an 'x' if the country of this row belong to this category. 

One of the missing value for ISO-alpha2 codes belongs to Namibia because its abbreviation `NA` is interpreted by R as a missing value!

The other missing values for ISO-alpha2 and ISO-alpha3 is related to [Sark](https://www.sark.co.uk/), which is "recognized by the United Nations Statistics Division (UNSD) as a separate territory" but was not accepted by ISO now for more than 20 years [@mccarthy-2020]. Recently a new application (see [PDF](https://www.sarkid.org/assets/pdf/SarkID%20Identity%20info%20v1_2.pdf)) will change that but currently Sark is still waiting for [ISO 3166 codes](https://www.iso.org/iso-3166-country-codes.html).


**Tab "clean"**: Recoding columns "LDC", "LLDC" and "SIDS" with 1 and 0 (1 = yes, belongs to this category, 0 = no, does not belong to this category) reduce most of their missing values. I have also recoded "Namibia" to repair their "NA" value.

**Tab "Region", "Sub-Region" and "Intermediate Region"**: One missing value in these regional categories is related to Antarctica which is not seen by the M49 scheme as a separated region. It has therefore no regional codes and names with the exception of the overall comprising global region. But it has M49 as well ISO-alpha codes.


:::::{.my-procedure}
:::{.my-procedure-header}
:::::: {#prp-02-clean-unsd-m49-data}
: Cleaning the UNSD M49 data file
::::::
:::
::::{.my-procedure-container}
To clean the data I have taken the following recoding actions in the script for the "clean" tab in

- Remove the global codes and names because they a redundant: All rows have global code "001" ("World").
- Rename the columns to get shorter names.
- Remove Antarctica because it is not seen as separate country.
- Replace `NA` in the column ISO-alpha2 Code" of Namibia with the string "NA".
- Recode the columns LDC, LLDC and SIDS with 0 and 1.
- Relocate the column "country" (previously "Country or Area") to the first column because than it easier to find some relevant content
- Sort the data alphabetically by "country".
::::
:::::






##### Summary

The UN geoscheme classification knows 248 countries divided the world into 

-  5 continents (Africa, Americas, Asia, Europe and Oceania): **Tab Region** 
- 17 sub-regions **Tab Sub-region**
-  9 intermediate regions focussing on South/Central America and Africa, otherwise not consistent **Tab Intermediate**
- 23 areas divides the continents into different parts **Tab Intermediate2**


## Combining M49 and WHR

To use the different region with the WHR data I will combine the M49 classification data scheme with the WHR data. Ideally I would combine the two datasets with their unique ISO3 code, but unfortunately the WHR data are lacking these data. Thew next best approach is to link the two datasets via the country names. But here I have to take into account that the authoritative country names are provided by M49 and not by the WHR data.

### Combine with country names

::: {.my-code-collection}
:::: {.my-code-collection-header}
::::: {.my-code-collection-icon}
:::::
:::::: {#exm-02-combine-whr-with-m49}
: Combine WHR data with M49 geoscheme
::::::

::::

::::{.my-code-collection-container}

::: {.panel-tabset}

###### Combine

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-02-combine-whr-m49}
: Combine WHR with UNSD M49 data
::::::
:::
::::{.my-r-code-container}
```{r}
#| label: combine-whr-m49

whr_2011_2024 <-  base::readRDS(
  "data/whr-cantril/rds/whr_2011_2024_arrange.rds")

unsd_class_finale <-  base::readRDS(
  "data/country-class/unsd/rds/unsd_class_finale.rds")

## left join: WHR as priority ########
whr_m49_first_try <- dplyr::left_join(
  x = whr_2011_2024,
  y = unsd_class_finale,
  by = dplyr::join_by(`Country name` == country),
  relationship = "many-to-many") |> 
  dplyr::relocate(iso3, .after = `Country name`)

## select columns and show distinct countries
 whr_m49_first_try |> 
    dplyr::select(
      `Country name`, iso2, iso3, m49
    ) |> 
    dplyr::distinct() |> 
    DT::datatable(class = 'cell-border compact stripe', 
                  options = list(
                    pageLength = 25,
                    lengthMenu = c(25, 50, 100, 200)
                    )
              )
```

::::
:::::


###### Show Missing

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-02-show-missing-iso3-codes}
: Show WHR countries with missing ISO3 codes
::::::
:::
::::{.my-r-code-container}
```{r}
#| label: show-missing-iso3-codes

whr_m49_first_try |> 
  dplyr::select(
      `Country name`, iso2, iso3, m49
    ) |> 
    dplyr::distinct() |> 
  dplyr::filter(base::is.na(iso3)) |> 
  DT::datatable(class = 'cell-border compact stripe', 
                  options = list(
                    pageLength = 25,
                    lengthMenu = c(5, 10, 15, 25)
                    )
              )
```

::::
:::::

:::

::::
:::::


### Differences WHR and M49


Combining the two datasets of M49 classification WHR leaves 16 country names in WHR empty because they do not match. The reasons are two fold:

1. Many names do not match because of different names even both dataset mean the same countries.
2. Some country names of the WHR data do not correspond to the official country classification.

The following table details the differences between the country names in the WHR and UNSD M49 datasets.

| WHR country name         | Official UNSD M49 country name                       | Comment                                                 |
|:--------------------------|:------------------------------------------------------|:--------------------------------------------------------|
| Bolivia                  | Bolivia (Plurinational State of)                     |                                                         |
| DR Congo                 | Democratic Republic of the Congo                     |                                                         |
| Hong Kong SAR of China   | China, Hong Kong Special Administrative Region       |                                                         |
| Iran                     | Iran (Islamic Republic of)                           |                                                         |
| Kosovo                   |                                                      | Not every country recognizes Kosovo (prov. code XKX)    |
| Lao PDR                  | Lao People's Democratic Republic                     |                                                         |
| Macedonia                | North Macedonia                                      | Renamed 2019, only wrong for the year 2019                                           |
| North Cyprus             |                                                      | Recognized only by Türkiye                              |
| Somaliland               |                                                      | Not recognized, self-declared independence from Somalia |
| Swaziland                | Eswatini                                             | Renamed 2018                                           |
| Syria                    | Syrian Arab Republic                                 |                                                         |
| Taiwan Province of China |                                                      | China -- as one of the world's largest and most influential countries -- considers Taiwan as part of its territory         |
| Tanzania                 | United Republic of Tanzania                          |                                                         |
| United Kingdom           | United Kingdom of Great Britain and Northern Ireland |                                                         |
| United States            | United States of America                             |                                                         |
| Venezuela                | Venezuela (Bolivian Republic of)                     |                                                         |

: Differences between WHR and UNSD M49 {#tbl-02-differences-whr-m49} {.striped .hover}

### WHR with revised country name

I change the WHR country designations to their official m49 country names. The exception are Kosovo, North Cyprus, Somaliland and Taiwan where no ISO codes exist.  These four names are set in angle brackets "[]" and remain without ISO and M49 codes.

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-02-revise-country-names}
: WHR data with official country names
::::::
:::
::::{.my-r-code-container}
```{r}
#| label: revise-country-names
#| results: hold


whr_2011_2024 <-  base::readRDS(
  "data/whr-cantril/rds/whr_2011_2024_arrange.rds")

## revise whr data with m49 names ################
whr_2011_2024_revised <-  whr_2011_2024 |> 
  dplyr::mutate(`Official country name (M49)` =
    dplyr::case_when(
      `Country name` == "Bolivia" ~ "Bolivia (Plurinational State of)",
      `Country name` == "DR Congo" ~ "Democratic Republic of the Congo",
      `Country name` == "Hong Kong SAR of China" ~ "China, Hong Kong Special Administrative Region",
      `Country name` == "Iran" ~ "Iran (Islamic Republic of)",
      `Country name` == "Kosovo" ~ "[Kosovo]",
      `Country name` == "Lao PDR" ~ "Lao People's Democratic Republic",
      `Country name` == "Macedonia" ~ "North Macedonia",
      `Country name` == "North Cyprus" ~ "[North Cyprus]",
      `Country name` == "Somaliland" ~ "[Somaliland]",
      `Country name` == "Swaziland" ~ "Eswatini",
      `Country name` == "Syria" ~ "Syrian Arab Republic",
      `Country name` == "Taiwan Province of China" ~ "[Taiwan Province of China]",
      `Country name` == "Tanzania" ~ "United Republic of Tanzania",
      `Country name` == "United Kingdom" ~ "United Kingdom of Great Britain and Northern Ireland",
      `Country name` == "United States" ~ "United States of America",
      `Country name` == "Venezuela" ~ "Venezuela (Bolivian Republic of)",
      TRUE ~ `Country name`
    )
  ) |> 
  dplyr::rename(`Country name (WHR)` = `Country name`) |> 
  dplyr::relocate(`Official country name (M49)`, .after = `Country name (WHR)`)

## select unique country rows  ################
whr_2011_2024_revised |> 
  dplyr::select(`Country name (WHR)`, `Official country name (M49)`) |> 
  dplyr::distinct() |> 
  DT::datatable(class = 'cell-border compact stripe', 
                  options = list(
                    pageLength = 25,
                    lengthMenu = c(25, 50, 100, 200)
                    )
              )

## save revised data as .rds file ################
my_save_data_file(
  "whr-cantril/rds", 
  whr_2011_2024_revised, 
  "whr_2011_2024_revised.rds")
```

::::
:::::

I have saved the WHR data file as `whr_2011_2024_revised.rds` under "data/whr_cantril/rds" with the new column `Official country name (M49)` and have changed the old WHR column `Country name` to `Country name (WHR)`.


### Combine WHR and M49 revised 

I am now in the position to combine the revised WHR file with the UNSD M49 classification file. I will show only the interesting columns but save the full new dataframe as `whr_m49` under the folder "data/whr_cantril/rds".

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-02-combine-whr-m49-revised}
: Combine WHR and M49 with revised country column
::::::
:::
::::{.my-r-code-container}
```{r}
#| label: combine-whr-m49-revised
#| results: hold

whr_2011_2024_revised <-  base::readRDS(
  "data/whr-cantril/rds/whr_2011_2024_revised.rds")

unsd_class_finale <-  base::readRDS(
  "data/country-class/unsd/rds/unsd_class_finale.rds")

## left join: WHR as priority ########
whr_m49 <- dplyr::left_join(
  x = whr_2011_2024_revised,
  y = unsd_class_finale,
  by = dplyr::join_by(`Official country name (M49)` == country),
      relationship = "many-to-many") 

whr_m49 |> 
  dplyr::select(
    `Country name (WHR)`, 
    `Official country name (M49)`,
    m49, iso2, iso3
    ) |> 
  dplyr::distinct() |> 
  DT::datatable(class = 'cell-border compact stripe', 
                  options = list(
                    pageLength = 25,
                    lengthMenu = c(25, 50, 100, 200)
                    )
              )

## save revised data as .rds file ################
my_save_data_file(
  "whr-cantril/rds", 
  whr_m49, 
  "whr_m49.rds")
```

::::
:::::




## Glossary

(Some of the abbreviation have at their end an additional "x" that is not part of the abbreviation. I chose this work around to distinguish these abbreviations from the same text chunks in one of the glossary entries. This is a bug in the {**glossary**} package.)

```{r}
#| label: glossary-table
#| echo: false

glossary_table()
```

------------------------------------------------------------------------

## Session Info {.unnumbered}

::::: my-r-code
::: my-r-code-header
Session Info
:::

::: my-r-code-container
```{r}
#| label: session-info

sessioninfo::session_info()
```
:::
:::::
