---
execute:
  cache: true
---

# Ranking 2024 {#sec-ranking-2024}


```{r}
#| label: setup
#| results: hold
#| include: false

base::source(file = paste0(here::here(), "/R/helper.R"))
```

::::: {#obj-chapter-template}
:::: {.my-objectives}
::: {.my-objectives-header}
Objective of this chapter
:::

::: {.my-objectives-container}
I will display the ranking data for the `r glossary("cantril ladder")` question of `r glossary("WHR")`data 2024. In contrast to figure 2.1 of the World Happiness Report 2025 I will use a `r glossary("lollipop plot")` instead a bar plot. Additionally I will only show the value for the cantril ladder without the breakdown into its causal components.

:::
::::
:::::


## Prepare data

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-02-prepare-data}
: Prepare 2024 ranking data
::::::
:::
::::{.my-r-code-container}
```{r}
#| label: prepare-data

ranking_2024 <- base::readRDS("data/whr-cantril/rds/whr_2011_2024.rds") |> 
    dplyr::filter(Year == 2024) |> 
    dplyr::arrange(Rank) |> 
    dplyr::mutate(`Country label` = 
        base::paste0(Rank, ". ", `Country name`, " (", `Ladder score`, ")")
        )
```

<center>(*For this R code chunk is no output available*)</center>
::::
:::::

## Cantril ladder ranking

```{r}
#| label: cantril-ranking
#| results: hold
#| include: false

cantril_ranking <-  function(df) {
    ggplot2::ggplot(df,
                ggplot2::aes(
                    x = forcats::fct_reorder(`Country label`, `Ladder score`) ,
                    y = `Ladder score`
                )) +
    ggplot2::geom_segment(ggplot2::aes(
        x = `Country label`,
        xend = `Country label`,
        y = 1,
        yend = `Ladder score`
    ), color = "blue") +
    ggplot2::geom_point(color = "darkblue", size = 4) +
    ggplot2::theme_light() +
    ggplot2::theme(
        panel.grid.major.x = ggplot2::element_blank(),
        panel.border = ggplot2::element_blank(),
        axis.ticks.x = ggplot2::element_blank(),
        axis.text.y = ggplot2::element_text(size = 20, hjust = 0),
        axis.text.x = ggplot2::element_text(size = 20),
        axis.title.x = ggplot2::element_text(size = 20, face = 'bold')
    ) +
    ggplot2::coord_flip() +
    ggplot2::xlab("") +
    ggplot2::ylab("Value of life evaluation (3 year average)")
}
```


::: {.my-code-collection}
:::: {.my-code-collection-header}
::::: {.my-code-collection-icon}
:::::
:::::: {#exm-show-ranking-2024}
: Show value of Cantril ladder by rank
::::::

::::

::::{.my-code-collection-container}

::: {.panel-tabset}

###### First part

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-show-ranking-2024-first-part}
: Show value of Cantril ladder by rank (first part)
::::::
:::
::::{.my-r-code-container}
```{r}
#| label: show-ranking-2024-first-part
#| fig-height: 20
#| fig-width: 12

cantril_ranking(dplyr::slice(ranking_2024, 1:50))
```

::::
:::::


###### Middle part

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-show-ranking-2024-middle-part}
: Show value of Cantril ladder by rank (middle part)
::::::
:::
::::{.my-r-code-container}
```{r}
#| label: show-ranking-2024-middle-part
#| fig-height: 20
#| fig-width: 12

cantril_ranking(dplyr::slice(ranking_2024, 51:100))

```

::::
:::::


###### Last part

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-show-ranking-2024-middle-part}
: Show value of Cantril ladder by rank (last part)
::::::
:::
::::{.my-r-code-container}
```{r}
#| label: show-ranking-2024-last-part
#| fig-height: 20
#| fig-width: 12

cantril_ranking(dplyr::slice(ranking_2024, 101:147))
```

::::
:::::

###### All together

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-02-show-ranking-2024-all-together}
: Show value of Cantril ladder by rank
::::::
:::
::::{.my-r-code-container}
```{r}
#| label: show-ranking-2024-all-together
#| fig-height: 60
#| fig-width: 12


cantril_ranking(ranking_2024)
```

::::
:::::

:::

::::
:::::




## Glossary

```{r}
#| label: glossary-table
#| echo: false

glossary_table()
```

------------------------------------------------------------------------

## Session Info

::::: my-r-code
::: my-r-code-header
Session Info
:::

::: my-r-code-container
```{r}
#| label: session-info

sessioninfo::session_info()
```
:::
:::::


