# Well-Being {#sec-well-being}

```{r}
#| label: setup
#| results: hold
#| include: false

base::source(file = paste0(here::here(), "/R/helper.R"))

ggplot2::theme_set(ggplot2::theme_bw())
```

:::::: {#obj-chapter-template}
::::: my-objectives
::: my-objectives-header
My objective for this chapter
:::

::: my-objectives-container
In this chapter, I will take a first take on the well-being ranking data
provided by the different [World Happiness Reports
2012-2024](https://worldhappiness.report/analysis/)
[@whr-download-analysis].

There are figures for all countries trajectories during the years. This
also includes Austria [@helliwell2024a, p.10]. My general aim is to
provide a comparison of the series of rankings for Austria with other
selected countries. I will focus on the so-called Cantril ladder of
subjective life evaluation [@nilsson2024].

As far as I can see from my first inspection, there are at least three
challenges:

1.  **Compatibility**: Check if all the data are compatible.
2.  **Completeness**: Check if the data for the time series are
    complete. I noticed already that there is no Excel data file
    for 2013. Bu there data in a PDF file. How to get these data?
    Another issue are the number of observed countries by year. How to
    get consistent rankings if the number of countries varies?
3.  **Visualization** Choose an appropriate visualization of the data.
    There are already up-to-date (2005-2023) visualization of
    trajectories. But these plots are visualizing the absolute value of
    the life evaluations country-by-country as the following example of
    Austria shows.

![Series of the values of subjective life evaluations for Austria
(2005-2023)](img/09-austria-well-being-trajectory-min.png){#fig-09-life-evaluation-austria
fig-alt="alt-text" fig-align="center" width="80%"}


@fig-09-life-evaluation-austria is a valuable figure to compare the development of the Austrian life evaluation index. But it is not useful for a comparison with other countries. Here I am thinking of two different visualizations:

1. Series of several selected countries in one figure so that it facilitates a comparison. In this case I will develop a graph with different colored lines for different countries.
2. With another figure I will try to capture the changes in rankings of some selected countries. The development of the absolute values of the life evaluations does not necessarily reflect the progress inside  countries because they are also influenced by global developments (e.g., COVID-19).
:::
:::::
::::::

## Import data

The first step is to find the life-evaluation data and to store the file for further R processing. I will store the files in my "whr" folder inside my "data" folder. (As I do not the chapter number of the final publication I am not going to use "chapXX" ("XX" standing for the chapter number) but "whr" as my folder name.)

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-create-whr-folder}
: Create folder for data files of the World Happiness Reports
::::::
:::
::::{.my-r-code-container}


The function `create_folder()` checks if the folder already exists. If this is the case it leaves it untouched, otherwise it creates the folder at the desired path.

```{r}
#| label: create-whr-folder

create_folder(base::paste0(here::here(), "/data/"))
create_folder(base::paste0(here::here(), "/data/whr"))
```

(*For this R code chunk is no output available*)


::::
:::::

The next code chunks combines two actions for each year:

1. At first I had to get the URL manually. Starting from the [WHR-Analysis website](https://worldhappiness.report/analysis/) I had to open each year and to look for the "Appendices & data" link. There I had to look for a data file.
2. After downloading all data file as Excel files I had to inspect them manually to find the sheet with the Cantril ladder data. I had to note the sheet number for the appropriate download. (Default is sheet 1). Additionally I had to note the file format "xls" or "xlsx" for each year to call the correct function `readxl::read_xls` or `readxl::read_xlsx`. (I refrained from the format guessing with the more general `readxl::read_exel()` function.) Furthermore I had to skip the first three comment lines in the 2015 dataset.

I am using the utility package [downloader](https://cran.r-project.org/web/packages/downloader/index.html) by Winston Chang because it provides a wrapper for the `utils::download.file()` function with the advantage that the download procedure works on Windows, Mac OS X, and other Unix-like platforms.


:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-download-whr-data}
: Download Data for the World Happiness Reports (WHR)
::::::
:::
::::{.my-r-code-container}


**Run the code just once manually**. It will download the data from the World Happiness Report (WHR) website for every year where data are available.


```{r}
#| label: download-whr-data
#| eval: false

## 2024 #########################################
downloader::download(
    url = "https://happiness-report.s3.amazonaws.com/2024/DataForFigure2.1+with+sub+bars+2024.xls",
    destfile = base::paste0(here::here(), "/data/whr/WHR-2024.xls")
)
WHR_data_2024 <- readxl::read_xls(base::paste0(here::here(), "/data/whr/WHR-2024.xls"))
save_data_file("whr", WHR_data_2024, "WHR_data_2024.rds")

## 2023 #########################################
downloader::download(
    url = "https://happiness-report.s3.amazonaws.com/2023/DataForFigure2.1WHR2023.xls",
    destfile = base::paste0(here::here(), "/data/whr/WHR-2023.xls")
)
WHR_data_2023 <- readxl::read_xls(base::paste0(here::here(), "/data/whr/WHR-2023.xls"))
save_data_file("whr", WHR_data_2023, "WHR_data_2023.rds")

## 2022 #########################################
downloader::download(
    url = "https://happiness-report.s3.amazonaws.com/2022/Appendix_2_Data_for_Figure_2.1.xls",
    destfile = base::paste0(here::here(), "/data/whr/WHR-2022.xls")
)
WHR_data_2022 <- readxl::read_xls(base::paste0(here::here(), "/data/whr/WHR-2022.xls"))
save_data_file("whr", WHR_data_2022, "WHR_data_2022.rds")

## 2021 #########################################
downloader::download(
    url = "https://happiness-report.s3.amazonaws.com/2021/DataForFigure2.1WHR2021C2.xls",
    destfile = base::paste0(here::here(), "/data/whr/WHR-2021.xls")
)
WHR_data_2021 <- readxl::read_xls(base::paste0(here::here(), "/data/whr/WHR-2021.xls"))
save_data_file("whr", WHR_data_2021, "WHR_data_2021.rds")


# 2020 #########################################
downloader::download(
    url = "https://happiness-report.s3.amazonaws.com/2020/WHR20_DataForFigure2.1.xls",
    destfile = base::paste0(here::here(), "/data/whr/WHR-2020.xls")
)
WHR_data_2020 <- readxl::read_xls(base::paste0(here::here(), "/data/whr/WHR-2020.xls"))
save_data_file("whr", WHR_data_2020, "WHR_data_2020.rds")


# 2019 #########################################
downloader::download(
    url = "https://s3.amazonaws.com/happiness-report/2019/Chapter2OnlineData.xls",
    destfile = base::paste0(here::here(), "/data/whr/WHR-2019.xls")
)
WHR_data_2019 <- readxl::read_xls(
    path = base::paste0(here::here(), "/data/whr/WHR-2019.xls"), 
    sheet = 2)
save_data_file("whr", WHR_data_2019, "WHR_data_2019.rds")


# 2018 #########################################
downloader::download(
    url = "https://s3.amazonaws.com/happiness-report/2018/WHR2018Chapter2OnlineData.xls",
    destfile = base::paste0(here::here(), "/data/whr/WHR-2018.xls")
)
WHR_data_2018 <- readxl::read_xls(
    path = base::paste0(here::here(), "/data/whr/WHR-2018.xls"), 
    sheet = 2)
save_data_file("whr", WHR_data_2018, "WHR_data_2018.rds")


# 2017 #########################################
downloader::download(
    url = "https://s3.amazonaws.com/happiness-report/2017/online-data-chapter-2-whr-2017.xlsx",
    destfile = base::paste0(here::here(), "/data/whr/WHR-2017.xlsx")
)
WHR_data_2017 <- readxl::read_xlsx(
    path = base::paste0(here::here(), "/data/whr/WHR-2017.xlsx"), 
    sheet = 3)
save_data_file("whr", WHR_data_2017, "WHR_data_2017.rds")


# 2016 #########################################
downloader::download(
    url = "https://s3.amazonaws.com/happiness-report/2016/Online-data-for-chapter-2-whr-2016.xlsx",
    destfile = base::paste0(here::here(), "/data/whr/WHR-2016.xlsx")
)
WHR_data_2016 <- readxl::read_xlsx(
    path = base::paste0(here::here(), "/data/whr/WHR-2016.xlsx"), 
    sheet = 3)
save_data_file("whr", WHR_data_2016, "WHR_data_2016.rds")

# 2015 #########################################
downloader::download(
    url = "https://s3.amazonaws.com/happiness-report/2015/Chapter2OnlineData_Expanded-with-Trust-and-Governance.xlsx",
    destfile = base::paste0(here::here(), "/data/whr/WHR-2015.xlsx")
)
WHR_data_2015 <- readxl::read_xlsx(
    path = base::paste0(here::here(), "/data/whr/WHR-2015.xlsx"), 
    sheet = 2,
    skip = 3)
save_data_file("whr", WHR_data_2015, "WHR_data_2015.rds")

## 2014 missing #######################################



## 2013 data in PDF ##################################
## store PDF file locally to prevent link rotting
downloader::download(
    url = "https://s3.amazonaws.com/happiness-report/2013/Chapter-2_online-appendix_9-5-13_final.pdf",
    destfile = base::paste0(here::here(), "/data/whr/WHR-PDF-2013.pdf")
)

## WHR_data_2013 will be generated and saved in the next section. 


# 2012 #########################################
downloader::download(
    url = "https://happiness-report.s3.amazonaws.com/2012/2012.xlsx",
    destfile = base::paste0(here::here(), "/data/whr/WHR-2012.xlsx")
)
WHR_data_2012 <- readxl::read_xlsx(
    path = base::paste0(here::here(), "/data/whr/WHR-2012.xlsx"), 
    sheet = 2)
save_data_file("whr", WHR_data_2012, "WHR_data_2012.rds")

```

(*For this R code chunk is no output available*)


::::
:::::

## Extract 2013 WHR Data from PDF

Data for the [World Happiness Report 2013](https://worldhappiness.report/ed/2013/#appendices-and-data) are not available in Excel format, unlike data from the other reports. Instead, the data are provided as a PDF file. (See also the second question of the [Frequently Asked Question](https://s3.amazonaws.com/happiness-report/2013/WHR-2013_FAQs_final.pdf))

This section extracts data I am interested in --- the survey measure of SWB (Subjective Well-being) from the Gallup World Poll (GWP) --- with the help of the `tabulapdf` package.

At first, I had to look for an appropriate R package because the one I had already used ([tabulizer](https://cran.r-project.org/web/packages/tabulizer/index.html)) was removed from CRAN. I found several packages for extracting PDF text (e.g., [pdftools](https://docs.ropensci.org/pdftools/), [PDE](https://cran.r-project.org/web/packages/PDE/vignettes/PDE.html), [PDFR](https://github.com/AllanCameron/PDFR)), but only [tabulapdf](https://docs.ropensci.org/tabulapdf/) is specialized in extracting table data.

I read the [introductory vignette](https://docs.ropensci.org/tabulapdf/articles/tabulapdf.html) completely because only in the final section I found a solution for my specific PDF. To understand the descriptions of the problems I had to master have a look at [Chapter 2 Appendix](https://s3.amazonaws.com/happiness-report/2013/Chapter-2_online-appendix_9-5-13_final.pdf). We need the cantril ladder data presented in "Table A2. National Averages in 2010-12". To extract the data I had to overcome three separate problems:

1. Table A2 spans over several pages (1-6).
2. On page 6 starts already the next table "Table A3. National Averages in 2005-07" which has to be excluded.
3. Pages 2-6 do not have column names.
4. Some of the column names in page 1 span two lines

The difficulties 1 and 2 arise because `tabulapdf` works page-oriented. Difficulty 3 originates from a faulty table design in the PDF. Difficulty 4 is a common problem that can be solved by setting the `col_names` argument to `FALSE`.

The solution I found is an [interactive table extraction](https://docs.ropensci.org/tabulapdf/articles/tabulapdf.html#interactive-table-extraction) for each page with the `tabulapdf::locate_area()` function and to select only the figures without columns names as demonstrated in the [use case section](https://docs.ropensci.org/tabulapdf/articles/tabulapdf.html#use-case-covid-19-treatments-in-italy) of the vignette.


:::::{.my-example}
:::{.my-example-header}
:::::: {#exm-extract-whr-table-data-2013}
: Extract WHR 2013 Table Data from PDF and Tidy Data
::::::
:::
::::{.my-example-container}

::: {.panel-tabset}

###### Extract

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-extract-pdf-table}
: Extract 2013 Table Data from PDF
::::::
:::
::::{.my-r-code-container}

**Run the code just once manually**. It will download the WHR 2013 PDF, generate interactively the table coordinates and save them for further processing.

```{r}
#| label: extract-pdf-table
#| eval: false

library(tabulapdf)

# set Java memory limit to 600 MB (optional)
options(java.parameters = "-Xmx600m")

# get the remote file and copy it to a temporary directory internally handled by R
f <- "https://s3.amazonaws.com/happiness-report/2013/Chapter-2_online-appendix_9-5-13_final.pdf"

# manual interactivity selection
# appropriate page will appear in the RStudio viewer
# point and drag to locate the areas coordinates
# see details in screenshot for interactive table extraction
# https://docs.ropensci.org/tabulapdf/articles/tabulapdf.html#interactive-table-extraction

page1 <- locate_areas(f, pages = 1)[[1]]
page2_5 <- locate_areas(f, pages = 2)[[1]]
page6 <- locate_areas(f, pages = 6)[[1]]


## store areas coordinates of WHR 2013 PDF data
create_folder(base::paste0(here::here(), "/data/whr/2013-coord"))

save_data_file("whr/2013-coord", page1, "page1.rds")
save_data_file("whr/2013-coord", page2_5, "page2_5.rds")
save_data_file("whr/2013-coord", page6, "page6.rds")

```


::::
:::::


###### Tidy PDF Data

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-tidy-pdf-data}
: Tidy 2013 WHR Data
::::::
:::
::::{.my-r-code-container}

**Run the code just once manually**. It will restore the coordinates of the table data for the WHR 2013 PDF, tidy the dataset and save the result as R data object in `.rds` format.

```{r}
#| label: tidy-pdf-data
#| eval: false


page1 <- base::readRDS("data/whr/2013-coord/page1.rds")
page2_5 <- base::readRDS("data/whr/2013-coord/page2_5.rds")
page6 <- base::readRDS("data/whr/2013-coord/page6.rds")

table1 <- extract_tables(
    f,
    pages = 1, 
    guess = FALSE, 
    col_names = FALSE,
    area = list(page1)
)

table2_5 <- extract_tables(
    f,
    pages = c(2, 3, 4, 5), 
    guess = FALSE, 
    col_names = FALSE,
    area = list(page2_5)
)

table6 <- extract_tables(
    f,
    pages = 6, 
    guess = FALSE, 
    col_names = FALSE,
    area = list(page6)
)

WHR_data_2013 <- dplyr::bind_rows(table1[[1]], 
                 table2_5[[1]], table2_5[[2]], table2_5[[3]], table2_5[[4]], 
                 table6[[1]])

column_names = c("Country", "Region", "Ladder", "Social support",
                 "Freedom", "Corruption", "Donation", "Generosity",
                 "Positive affect", "Negative affect", "Happiness (Yesterday)",
                 "GDP per capita", "Healthy life expectancy")

WHR_data_2013 <- WHR_data_2013 |> 
    dplyr::rename_with(~column_names, dplyr::everything()) |> 

WHR_data_2013[17, 1] <- "Bosnia and Herzegovina"
WHR_data_2013[27, 1] <- "Central African Republic"
WHR_data_2013[34, 1] <- "Congo (Brazzaville)"
WHR_data_2013[43, 1] <- "Dominican Republic"
WHR_data_2013[146, 1] <- "Trinidad and Tobago"
WHR_data_2013[154, 1] <- "United Arab Emirates"

## missing data (data is in wrong previous NA row)
WHR_data_2013[154, 6] <- 0.355

WHR_data_2013 <- na.omit(WHR_data_2013)


save_data_file("whr", WHR_data_2013, "WHR_data_2013.rds")
```

(*For this R code chunk is no output available*)

::::
:::::

:::

::::
:::::

***





## Inspect Data

:::::{.my-example}
:::{.my-example-header}
:::::: {#exm-ID-text}
: Inspect WHR data that contains ladder score and ranking
::::::
:::
::::{.my-example-container}

::: {.panel-tabset}

###### 2024

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-glimpse-whr-data-2024}
: Glimpse at WHR 2024 ladder score data
::::::
:::
::::{.my-r-code-container}
::: {#lst-glimpse-whr-data-2024}
```{r}
#| label: glimpse-whr-data-2024

WHR_2024 <- base::readRDS("data/whr/WHR_data_2024.rds")
dplyr::glimpse(WHR_2024)

```
Glimpse at ladder score 2024
:::

::::
:::::


###### 2023

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-glimpse-whr-data-2023}
: Glimpse at WHR 2023 ladder score data
::::::
:::
::::{.my-r-code-container}
::: {#lst-glimpse-whr-data-2023}
```{r}
#| label: glimpse-whr-data-2023

WHR_2023 <- base::readRDS("data/whr/WHR_data_2023.rds")
dplyr::glimpse(WHR_2023)
```

Glimpse at ladder score 2023
:::

::::
:::::

###### 2022

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-glimpse-whr-data-2022}
: Glimpse at WHR 2022 ladder score data
::::::
:::
::::{.my-r-code-container}
::: {#lst-glimpse-whr-data-2022}
```{r}
#| label: glimpse-whr-data-2022

WHR_2022 <- base::readRDS("data/whr/WHR_data_2022.rds")
dplyr::glimpse(WHR_2022)
```

Glimpse at ladder score 2022
:::

::::
:::::

###### 2021

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-glimpse-whr-data-2021}
: Glimpse at WHR 2021 ladder score data
::::::
:::
::::{.my-r-code-container}
::: {#lst-glimpse-whr-data-2021}
```{r}
#| label: glimpse-whr-data-2021

WHR_2021 <- base::readRDS("data/whr/WHR_data_2021.rds")
dplyr::glimpse(WHR_2021)
```

Glimpse at ladder score 2021
:::

::::
:::::

###### 2020

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-glimpse-whr-data-2020}
: Glimpse at WHR 2020 ladder score data
::::::
:::
::::{.my-r-code-container}
::: {#lst-glimpse-whr-data-2020}
```{r}
#| label: glimpse-whr-data-2020

WHR_2020 <- base::readRDS("data/whr/WHR_data_2020.rds")
dplyr::glimpse(WHR_2020)
```

Glimpse at ladder score 2020
:::

::::
:::::

###### 2019

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-glimpse-whr-data-2019}
: Glimpse at WHR 2019 ladder score data
::::::
:::
::::{.my-r-code-container}
::: {#lst-glimpse-whr-data-2019}
```{r}
#| label: glimpse-whr-data-2019

WHR_2019 <- base::readRDS("data/whr/WHR_data_2019.rds")
dplyr::glimpse(WHR_2019)
```

Glimpse at ladder score 2019
:::

::::
:::::

###### 2018

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-glimpse-whr-data-2018}
: Glimpse at WHR 2018 ladder score data
::::::
:::
::::{.my-r-code-container}
::: {#lst-glimpse-whr-data-2018}
```{r}
#| label: glimpse-whr-data-2018

WHR_2018 <- base::readRDS("data/whr/WHR_data_2018.rds")
dplyr::glimpse(WHR_2018)
```

Glimpse at ladder score 2018
:::

::::
:::::

###### 2017

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-glimpse-whr-data-2017}
: Glimpse at WHR 2017 ladder score data
::::::
:::
::::{.my-r-code-container}
::: {#lst-glimpse-whr-data-2017}
```{r}
#| label: glimpse-whr-data-2017

WHR_2017 <- base::readRDS("data/whr/WHR_data_2017.rds")
dplyr::glimpse(WHR_2017)
```

Glimpse at ladder score 2017
:::

::::
:::::

###### 2016

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-glimpse-whr-data-2016}
: Glimpse at WHR 2016 ladder score data
::::::
:::
::::{.my-r-code-container}
::: {#lst-glimpse-whr-data-2016}
```{r}
#| label: glimpse-whr-data-2016

WHR_2016 <- base::readRDS("data/whr/WHR_data_2016.rds")
dplyr::glimpse(WHR_2016)
```

Glimpse at ladder score 2016
:::

::::
:::::


###### 2015

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-glimpse-whr-data-2015}
: Glimpse at WHR 2015 ladder score data
::::::
:::
::::{.my-r-code-container}
::: {#lst-glimpse-whr-data-2015}
```{r}
#| label: glimpse-whr-data-2015

WHR_2015 <- base::readRDS("data/whr/WHR_data_2015.rds")
dplyr::glimpse(WHR_2015)
```

Glimpse at ladder score 2015
:::

::::
:::::

###### 2013

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-glimpse-whr-data-2013}
: Glimpse at WHR 2013 ladder score data
::::::
:::
::::{.my-r-code-container}
::: {#lst-glimpse-whr-data-2013}
```{r}
#| label: glimpse-whr-data-2013

WHR_2013 <- base::readRDS("data/whr/WHR_data_2013.rds")
dplyr::glimpse(WHR_2013)
```

Glimpse at ladder score 2013
:::

::::
:::::


###### 2012

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-glimpse-whr-data-2012}
: Glimpse at WHR 2012 ladder score data
::::::
:::
::::{.my-r-code-container}
::: {#lst-glimpse-whr-data-2012}
```{r}
#| label: glimpse-whr-data-2012

WHR_2012 <- base::readRDS("data/whr/WHR_data_2012.rds")
dplyr::glimpse(WHR_2012)
```

Glimpse at ladder score 2012
:::

::::
:::::


:::

::::
:::::

Inspecting the datasets it turned out that there is some variety in the number of rows (observed countries), number of columns (variables) and column names. I am interested only in the country column and the "ladder score" (sometimes "happiness score" and in 2012 just "ladder").

