# Countries & Variables {#sec-summary-table}

```{r}
#| label: setup
#| results: hold
#| include: false

base::source(file = paste0(here::here(), "/R/helper.R"))

ggplot2::theme_set(ggplot2::theme_bw())
options(show.signif.stars = FALSE)

```

## Table of content {.unnumbered}

::: {#obj-chapter-template}
::: my-objectives
::: my-objectives-header
Inspect countries and variables of the Summary Table
:::

::: my-objectives-container
The idea of this chapter is to get an idea what kind of data is
available for different countries. Scrolling the Summary Table
interactively is a not very efficient way. Even filtering for countries
does sometimes not work (for instance: it works for Austria, but not for
Germany)[^81-countries-and-variables-1]

One idea that comes to my mind is a list of countries and the number of their rows as an indicator how many variables are available. Perhaps I will invent other filters as well for a better understanding of the data.
:::
:::
:::

[^81-countries-and-variables-1]: Maybe a bug in the program of the web
    interface?

## Data

### Download Data

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-81-download-data}
: Numbered R Code Title
::::::
:::
::::{.my-r-code-container}
```{r}
#| label: download-data
#| eval: false

## define variables
url = "https://wid.world/wp-content/uploads/2019/05/WID_SummaryTable_24May2019.xlsx"
chapter_folder = "chap81"
file_name = "summary_table.xlsx"


## create folder for chapter if not already done
chap_folder <-
        base::paste0(
            here::here(),
            paste0("/data/", chapter_folder, "/")
        )
    if (!base::file.exists(chap_folder))
    {base::dir.create(chap_folder)}

## get summary table
destfile <- base::paste0(chap_folder, file_name)
utils::download.file(url, destfile)
summary_table <- readxl::read_xlsx(destfile)


## save cleaned data
save_data_file("chap81", summary_table, "summary_table.rds")

```



(*For this R code chunk is no output available*)
::::
:::::

### Explore data

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-81-explore-data}
: Explore Data of the Summary Table
::::::
:::
::::{.my-r-code-container}
```{r}
#| label: explore-data
#| results: hold

summary_table <- readRDS("data/chap81/summary_table.rds")

skimr::skim(summary_table)
```





::::
:::::

## Inspect Data

### Number of Variables per Country

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-81-list-variable-numbers-per-country}
: Number of Records (=Variables) per Country
::::::
:::
::::{.my-r-code-container}
```{r}
#| label: list-variable-numbers-per-country

summary_table |> 
    dplyr::count(Country, `Country Code`) |> 
    DT::datatable()

```


::::
:::::

### Full Table

Another interesting type of table for inspecting data is an interactive table with all columns, except the very long descriptions of the columns 13-16. To get a better view I have created a special page with a full-page layout covering the full width of the browser: Go to @sec-full-summary-table.

### List Variables of the Summary Table

Another important type of table for understanding the datasets is a table with the short name of the variables with the simple, technical and methodical description.

:::::{.my-r-code}
:::{.my-r-code-header}
:::::: {#cnj-81-unique-variables}
: Unique `Short name of variables`
::::::
:::
::::{.my-r-code-container}
```{r}
#| label: unique variables

summary_table |>
    dplyr::distinct(`Short name of variable`)
```


::::
:::::

### Explain All 184 Variables of the Summary Table

It turned out that the number of unique 'Short name of variable` is the same number of rows as for the country with the most rows (= USA with 184 rows). It is therefore a reasonable assumption that all unique variable names can be found in the USA part of the database. See @sec-explain-variables.

### Explain Variables for Austria of the Summary Table

I believe that I will not need most of the 184 variables explained in the Summary Table. As a first approach I could start with the 21 variables that are registered with Austria. See @sec-explain-austria-variables.

::: {.callout-caution #cau-quadruplicate}
###### Post- & Pre-tax national income occur quadruple

It turned out that two variables (Post-tax national income and Pre-tax national income) occur four times. But this happen the case also for other countries, like USA, Germany, Belgium etc.

Maybe these variables were added with different research projects?

:::

So in the end there are not 184 variables in total but 178 and Austria has only 15 variable!
